Things to run after normalization
========================================================
  
## Libraries
```{r}
setwd("/big_data/redgar/BECon")
library(reshape)
library(ggplot2)
library(RColorBrewer)
library(mclust)
library(IlluminaHumanMethylation450k.db)
```

## Load combatted, cell adjusted data with no replicates
```{r}
load("cell_adjusted_Mcombat_Bmiq_BLBR_Beta_noreplicates.RData") #441198     66
## need cell comp adjusted data and need to remove replicates (this is adjusted with seperated brain regions)
```

#### Any meta data
```{r}
meta<-read.csv("SampleInfo.csv")
fix_names<-unlist(lapply(meta$X, function(x) gsub(" ","", x , fixed=TRUE)))
meta$X<-fix_names
meta<-meta[which(meta$X%in%colnames(combat_BLBR_Beta)),]
meta<-meta[match(colnames(combat_BLBR_Beta), meta$X),]
```

## One sample (out of 17 individuals) will be remived as it is missing the blood sample
## NOTE CODE SAYS PBMC BUT IT IS WHOLE BLOOD
```{r}
PBMC_indx<-which(meta$TissueType=="PBMC") #16
BRAIN7_indx<-which(meta$TissueType=="BRAIN7") #17
BRAIN10_indx<-which(meta$TissueType=="BRAIN10") #17
BRAIN20_indx<-which(meta$TissueType=="BRAIN20") #16

    #which PBMC is missing [individual 173]
    meta[BRAIN7_indx,][which(!(meta[BRAIN7_indx,]$SubjectNumC%in%meta[PBMC_indx,]$SubjectNumC)),]
    #which BRAIN20 is missing [individual 221]
    meta[BRAIN7_indx,][which(!(meta[BRAIN7_indx,]$SubjectNumC%in%meta[BRAIN20_indx,]$SubjectNumC)),]

## can work without one brain 20 but sample without blood is not useable
combat_BLBR_Beta<-combat_BLBR_Beta[,which(!(colnames(combat_BLBR_Beta)%in%c("BA20173","BA10173","BA7173")))] #441198     65
meta<-meta[which(meta$X%in%colnames(combat_BLBR_Beta)),]
meta<-meta[match(colnames(combat_BLBR_Beta), meta$X),]


## Need to add 221 BRAIN20 for cor (just as NA though)
combat_BLBR_Beta_correlation<-as.data.frame(combat_BLBR_Beta)
combat_BLBR_Beta_correlation$BA20221<-"NA"
BA20221<-data.frame(X="BA20221", SampleID="BA20221", SubjectNumC=221, Barcode="NA", Section="NA", Well="NA", TissueType="BRAIN20", Replicate="NA", Gender="NA", Age="NA",  PH="NA", RefrigerationDelayHrs="NA", Row="NA")
meta_correlation<-rbind(meta,BA20221)
meta_correlation<-meta_correlation[which(meta_correlation$X%in%colnames(combat_BLBR_Beta_correlation)),]
meta_correlation<-meta_correlation[match(colnames(combat_BLBR_Beta_correlation), meta_correlation$X),]

combat_BLBR_Beta<-combat_BLBR_Beta_correlation
save(combat_BLBR_Beta, file="cell_adjusted_Mcombat_Bmiq_BLBR_Beta_noreplicates_16_All.RData")
```


## Prep data for correlation of CpGs Between Blood and each brain Region
## actual correlations will be done in python
```{r}
PBMC_indx<-which(meta_correlation$TissueType=="PBMC") #16
BRAIN7_indx<-which(meta_correlation$TissueType=="BRAIN7") #16
BRAIN10_indx<-which(meta_correlation$TissueType=="BRAIN10") #16
BRAIN20_indx<-which(meta_correlation$TissueType=="BRAIN20") #16

PBMC_BLBR<-combat_BLBR_Beta_correlation[,PBMC_indx]
BRAIN7_BLBR<-combat_BLBR_Beta_correlation[,BRAIN7_indx]
BRAIN10_BLBR<-combat_BLBR_Beta_correlation[,BRAIN10_indx]
BRAIN20_BLBR<-combat_BLBR_Beta_correlation[,BRAIN20_indx]

# put in PBMC sample order
BRAIN7_BLBR<-BRAIN7_BLBR[,c(2,3,1,5,4,6,9,7,8,10,11,12,13,14,15,16)]
BRAIN10_BLBR<-BRAIN10_BLBR[,c(1,3,2,6,4,5,7,9,8,11,10,12,14,13,16,15)]
BRAIN20_BLBR<-BRAIN20_BLBR[,c(2,1,3,4,5,6,9,7,8,10,16,11,12,13,15,14)]

# for python
write.csv(meta, file="Python/Meta_BLBR_jan22.csv")
write.csv(PBMC_BLBR, file="Python/PBMC_BLBR_jan22.csv")
write.csv(BRAIN7_BLBR, file="Python/BRAIN7_BLBR_jan22.csv")
write.csv(BRAIN10_BLBR, file="Python/BRAIN10_BLBR_jan22.csv")
write.csv(BRAIN20_BLBR, file="Python/BRAIN20_BLBR_jan22.csv")
```


## PYTHON CORRELATION
                        
# Python correlations from python BLBR_Correlation_Permutation_jan22.py produces correlation_jan22.csv 
```{r}
load("cell_adjusted_Mcombat_Bmiq_BLBR_Beta_noreplicates.RData") #441198     66
combat_BLBR_Beta_adjusted<-combat_BLBR_Beta

correlations_BLBR<-read.csv("correlation_jan22.csv", sep="\t")
         
colnames(correlations_BLBR)<-c("CpG","BRAIN7","BRAIN10","BRAIN20")
correlations_BLBR$CpG<-rownames(combat_BLBR_Beta)

correlations_BLBR_melt<-melt(correlations_BLBR, id="CpG")
```


## Five correlations with shuffling PBMC order and therefore unpairing blood and brain (BLBR_Correlation_Permutation_jan22_permutations.py)
```{r}
correlations_BLBR_random1<-read.csv("Python/Rnd_correlations/correlation_jan22_permutation10.csv", sep="\t")
correlations_BLBR_random2<-read.csv("Python/Rnd_correlations/correlation_jan22_permutation9.csv", sep="\t")
correlations_BLBR_random3<-read.csv("Python/Rnd_correlations/correlation_jan22_permutation8.csv", sep="\t")
correlations_BLBR_random4<-read.csv("Python/Rnd_correlations/correlation_jan22_permutation7.csv", sep="\t")
correlations_BLBR_random5<-read.csv("Python/Rnd_correlations/correlation_jan22_permutation6.csv", sep="\t")
correlations_BLBR_random<-rbind(correlations_BLBR_random1, correlations_BLBR_random2,correlations_BLBR_random3,
                                correlations_BLBR_random4,correlations_BLBR_random5)
save(correlations_BLBR_random, file="correlation_jan22_permutations.RData")


load("correlation_jan22_permutations.RData")
colnames(correlations_BLBR_random)<-c("CpG","BRAIN7","BRAIN10", "BRAIN20")
correlations_BLBR_random_melt<-melt(correlations_BLBR_random, id="CpG")


#PLot Actual
## Correlation Histogram
ggplot(correlations_BLBR_melt, aes(value, fill=variable))+geom_histogram(colour = "black")+theme_bw()+
  facet_wrap(~variable, ncol=1)+scale_fill_manual(values=c("#fb6a4a","#ef3b2c","#cb181d"))

ggplot(correlations_BLBR_melt_unadjusted, aes(value, fill=variable))+geom_histogram(colour = "black")+theme_bw()+
  facet_wrap(~variable, ncol=1)+scale_fill_manual(values=c("#fb6a4a","#ef3b2c","#cb181d"))

ggplot(correlations_BLBR_random_melt, aes(value, fill=variable))+geom_histogram(colour = "black")+theme_bw()+
  facet_wrap(~variable, ncol=1)+scale_fill_manual(values=c("#fb6a4a","#ef3b2c","#cb181d"))


# Plot all together
correlations_BLBR_random_melt$Distribution<-"Random Permutations"
correlations_BLBR_melt$Distribution<-"Adjusted for Cell Type"

correlations_plot<-rbind(correlations_BLBR_melt,correlations_BLBR_random_melt)
correlations_plot$Distribution<-as.factor(correlations_plot$Distribution)
correlations_plot$Distribution<-factor(correlations_plot$Distribution, levels=c("Random Permutations","Adjusted for Cell Type"))

ggplot(correlations_plot, aes(value,  fill=interaction(Distribution,variable)))+
  geom_density(color="black", alpha=0.5)+theme_bw()+facet_wrap(~variable, ncol=1)+
  scale_fill_manual(values=c("grey","#fb6a4a","grey","#ef3b2c","grey","#cb181d"))+
  xlab("Correlation")
```


