Differential Methylation
========================================================
  
## Libraries
```{r}
suppressPackageStartupMessages(library(methylumi,wateRmelon, quietly = TRUE))
setwd("/big_data/redgar/BECon")
```


## make the methlumi objects
```{r}
allFile <- ("BLBR_alldat.txt") 
qcFile <- ("BLBR_QC.txt")
betaFile <- ("BLBR_avgbeta.txt") 

BLBR<- lumiMethyR(allFile)
BLBR.2 <- methylumiR(allFile, qcfile = allFile)
save(BLBR,BLBR.2, file="BLBR_methlylumi_second.RData")
```


The function *lumiMethyR()* coerces the object (**allFile**) into a **MethyLumiM class object**, which contains those four elements (exprs, methylated, unmethylated, and detection p value).

The function *methylumiR()* coerces the object (**betaFile**) into a **MethyLumiSet class object**, which holds the intensities and the beta values.  This is also where the sample information (the sampleFile) can be added as phenoData.


#### Pvalues of samples on each array

dont have for blbrGenomstudio run (could rerun)
```{r}
load("BLBR_methlylumi_second.RData")
avgPval <- colMeans(pvals(BLBR))
par(las = 2)

barplot(avgPval, ylab = "Average P-Value", col=meta$Sample_Plate) 

#p-values for detection for the samples
meta$Det_pval<-avgPval

ggplot(meta)+geom_boxplot(aes(as.factor(Barcode), Det_pval, fill=as.factor(Barcode)))+
  geom_point(aes(as.factor(Barcode), Det_pval, group=SampleID), shape=19, color="grey70",
             position = position_jitter(w = 0.25))+theme_bw()
```


# Normalization object
```{r}
BLBR<-BLBR.2
dim(BLBR) # 485577      70 
```


## Probe removal
## normalize


### F. Probe Filtering, Courtesy of yours truly, Sumaiya Islam 

##### Removal of SNP Probes
We remove the SNP probes as they are used as an internal control to ensure your samples are what you think they are and are not used for any methylation analysis.

```{r HTT_Leavitt_SNPprobes, echo=FALSE}
BLBR <- BLBR[substring(featureNames(BLBR),1,2) != "rs", ]
dim(BLBR) # probes = 485512, n = 70
```


##### Cross-hybridizing probes

Some probes have been found to cross-hybridize with other chromosomes (Price et al. 2013 *Epigenetics*).  It is at the discretion of the user whether or not they want to remove these cross-hybridizing probes, since it isn't a guarantee that they will cross-hybridize every time the experiment is run.  Probes that cross-hybridize to the sex chromosomes are typically removed, as they run a higher risk of confounding the data than probes that cross-hybridize to autosomal chromosomes.  

We remove probes which bind multiple locations in the genome as long as one of the locations is on the XY chromosome. The reason for this is as the X and Y chromosomes are not balanced amongst our samples (males vs females) we have to remove all probes which bind them so as to not skew the normalization. We do not remove multiple binders of probes which bind at various sites only in the autosomal chromosomal regions because they will most likely not skew our normalization and when we obtain our "hit list" we will verify them using pyrosequencing in order to determine if the specific site of interest is providing the signal we are seeing.

```{r}
xy_hit_index <- which(fData(BLBR)$XY_Hits == "XY_NO")
(n.XYcrosshybrid.probes<-(length(featureNames(BLBR))-length(xy_hit_index)))
BLBR <- BLBR[xy_hit_index, ] 
dim(BLBR) # probes = 473124, n = 70

auto_hit_index <- which(fData(BLBR)$Autosomal_Hits== "A_NO")
(n.autocrosshybrid.probes<-(length(featureNames(BLBR))-length(auto_hit_index)))
BLBR <- BLBR[auto_hit_index, ] 
dim(BLBR) # probes = 443575, n = 70
```

We have removed 42002 probes. This leaves us with 443575 probes for our analysis.


## going to removed these after all norm etc because didn't think to do it now, plus turns out they were interesting
```{r}
SNPCpG<-fData(BLBR)[,c(5,8)]
save(SNPCpG, file="SNPCpG.RData")
```




# Watermelon Filter and Normalization
# BMIQ (probe type normalization)


```{r}
# waterMelon Bad probe filtration
BLBR.pf<-pfilter(BLBR)

#0 samples having 1 % of sites with a detection p-value greater than 0.05 were removed 
#Samples removed:  
#1035 sites were removed as beadcount <3 in 5 % of samples 
#1671 sites having 1 % of samples with a detection p-value greater than 0.05 were removed 
dim(BLBR.pf) #  441198       70 


BLBR<-BLBR.pf
save(BLBR, file = "BLBR(fully_filtered).RData")
```

We have removed 44379 probes. This leaves us with 441198 probes for our analysis.


## beta distribution
```{r}
library(reshape)
library(ggplot2)
load("BLBR(fully_filtered).RData")
Beta_sample<-betas(BLBR)

meta<-read.csv("SampleInfo.csv")
fix_names<-unlist(lapply(meta$X, function(x) gsub(" ","", x , fixed=TRUE)))
meta$X<-fix_names
meta<-meta[which(meta$X%in%colnames(Beta_sample)),]
meta<-meta[match(colnames(Beta_sample), meta$X),]


#looking good? Take just a random sample of CpGs for plotting (otherwise so slow)
Beta_sample_melted<- melt(Beta_sample)
#remove NAs before plotting (otherwise get many non-inifnite warnings)
Beta_Plot<-Beta_sample_melted[which(!(is.na(Beta_sample_melted$value))),]
#add meta
Beta_Plot<-merge(Beta_Plot,meta, by.x="Var2", by.y="SampleID")

ggplot(Beta_Plot, aes(value, group=Var2, color=TissueType))+
  geom_density()+theme_bw()+ scale_color_manual(values=c("#fb6a4a","#ef3b2c","#cb181d","cornflowerblue"), name="Tissue")+
  xlab("DNAm Beta Value")+ylab("Density")

```


# BMIQ
```{r}
library(RPMM)
bmiq_BLBR<-BMIQ(BLBR) #small number to fit for test (defult is 50000, tested with nfit=100)

BLBR<-bmiq_BLBR
dim(BLBR) # 441198      70 
save(BLBR, file = "BLBR_prbFilter_BMIQnrm.RData")
```

